---
title: "CA and MCA of NYCABS dataset"
author: "Katerina Dimitrova, Jose Romero, Sergi Munoz"
date: "March 18, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r,include=FALSE}
setwd("C://Users/Sergi/Desktop/Sergi/ADEI") #Change 
#setwd("D:/ADEI/ADEI.git/trunk") #Change

```

# Load Required Packages: to be increased over the course
```{r}
rm(list=ls())
requiredPackages <- c("effects","FactoMineR","car", "factoextra","RColorBrewer","ggplot2","missMDA","mvoutlier","dplyr","ggmap","ggthemes","knitr","MVA")
missingPackages <- requiredPackages[!(requiredPackages %in% installed.packages()[,"Package"])]

if(length(missingPackages)) install.packages(missingPackages)
lapply(requiredPackages, require, character.only = TRUE)

# Useful function and Data
load("Taxi5000_raw_DataCleanv4.RData")
calcQ <- function(x) {
  s.x <- summary(x)
  iqr<-s.x[5]-s.x[2]
  list(souti=s.x[2]-3*iqr, mouti=s.x[2]-1.5*iqr, min=s.x[1], q1=s.x[2], q2=s.x[3],        q3=s.x[5], max=s.x[6], mouts=s.x[5]+1.5*iqr, souts=s.x[5]+3*iqr ) }
```



# Correspondence Analysis: f.cost (discretization of Total_amount) vs f.hour and f.tt and period
In this section we will interprete two Correspondence Analysis executed on our discretizated target variable (f.total) with 2 different factors of our data: pick_up_period and f.espeed.

## Total_amount vs pick_up_period
We generate the contingency table for this two variables and execute the chisq test.

```{r}
tt<-table(df[,c("f.total","pick_up_period")])
tt
prop.table(tt,1)
prop.table(table(df$pick_up_period))

```
Before getting into the chisq test, taking a look to the contingency table we can actually say very little, because the categories of pick_up_period are not well balanced and so "afternoon" is always taking a higher number of individuals for each category.

For the previous reason, we look at their marginal tables to better extrapolate our first impressions.
The marginal table relative to f.total it is just a confirmation for what we already said (not very helpfull): we have an unbalanced distribution: most of the trips were made during the afternoon in every range of price contended.


```{r}
prop.table(tt,2)
prop.table(table(df$f.total))
```

Then, moving on to the second marginal table (now, relative to pick_up_period variable) things get more even.
we can appreciate that the cheapest trips take place, in a greater proportion, during the morning, as me can also say that during the night occur the most expensive trips. On the other hand, the medium price categories are more associated to the afternoon.   

```{r}
chisq.test(tt)
```

From the chi-square test we see that the p.value is less than 0.05 which meand that we can reject the Nul hypotesis thus the two factors are dependent. The df is 9 which means that the chi value of the threshold is 16,9 and we see that we have X-squared = 36,42 which is bigger than 16,9. => We can deffinetly reject the hypotesis. Factors are dependent and so we can affirm that the conclusions made below are statistically significant.


### CA plot interpretation
The plot visually represents all the information we could have interpreted from the tables:

- At night the trips tend to be more expensive. 
- In the afternoon we find middle prices 
- Cheap trips stay between valley and morning.


```{r}
res.ca<-CA(tt)
lines(res.ca$row$coord[,1],res.ca$row$coord[,2],lwd=2,col="blue")
summary(res.ca,dig=2)
fviz_ca_biplot(res.ca,repel=TRUE)+theme_bw()
```


### Eigenvalues and dominant axes analysis. How many axes we have to consider
The two first dimensions explain 99% of the variance, so there is no need of further exploration. (Actually, only the first dim has more than 94% of the variance). 
We can be confident that the patterns we see in the CA plot represent the patterns that we would see if we could peer into n-dimensional space.
```{r}
res.ca$eig
fviz_eig(res.ca)

```


### Individuals
For the next output we deduce that the more extreme individuals are actually the most contributive ones for the specific axe of their coordinates.

```{r}
names(res.ca)
res.ca$row$contrib[,1:2]
res.ca$row$coord[,1:2]

```

## Total_amount vs f.ttime

```{r}
tt<-table(df[,c("f.total","f.ttime")])
tt
prop.table(tt,1)
prop.table(table(df$f.ttime))
chisq.test(tt)
prop.table(tt,2)
prop.table(table(df$f.total))

#From the plot we see that the longer the trips is the more it costs which was also expected.
res.ca<-CA(tt)
lines(res.ca$row$coord[,1],res.ca$row$coord[,2],lwd=2,col="blue")
#We need to consider the first 2 dimentions which together explain 90% of the data.
summary(res.ca)
res.ca$row$contrib[,1:2]
res.ca$row$coord[,1:2]
res.ca$row$cos2[,1:2]
#The 2st two eigenectors give us 90% of the variance so they are sufficient. 
res.ca$eig
sum(res.ca$eig[1:2,1])/sum(res.ca$eig[1:3,1])
sum(res.ca$eig[1:3,1])
res.ca$call$marge.col

fviz_eig(res.ca)
summary(res.ca,dig=2)
fviz_ca_biplot(res.ca,repel=TRUE)+theme_bw()

```

# MCA using multivariant

```{r}
names(df)
#explain we don't use this variable cause it is too bad balanced
summary(df$Store_and_fwd_flag)
#add espeed factor as active (we need to generate it)
vars_cat <- names(df)[c(1,19,20,22,27,29:35,39,41,42,43)]
#add f.totalamount as quali.sup
res.mca<-MCA(df[,c(vars_cat,"Total_amount")],quali.sup=c(4,13),quanti.sup=17)

# We see that all of the eigenvectors are really small, so we will use a similiar rule like in the PCA and take all EW with value >1 => the first 26 dimentions.
fviz_eig(res.mca)
fviz_screeplot(res.mca, addlabels = TRUE, ylim = c(0, 45))
get_eigenvalue(res.mca)
names(res.mca)
summary(res.mca,nbelements=50)
mean(res.mca$eig) 

# Individual Representation
plot.MCA(res.mca,choix=c("ind"),cex=0.8)
plot.MCA(res.mca,choix=c("ind"),invisible=c("ind"),cex=0.8)

# Representation of categories
plot.MCA(res.mca,choix=c("ind"),invisible=c("ind"),axes=c(1,2))
lines(res.mca$quali.sup$coord[3:4, 1],res.mca$quali.sup$coord[3:4,2],lwd=2,col="darkgreen") # Trencada dels nivells de f.fare_amount #??
lines(res.mca$quali.sup$coord[1:2,1],res.mca$quali.sup$coord[1:2,2],lwd=2,col="darkblue") # Trencada AnyTip
names(res.mca)

# Contributions of rows to dimension 1
fviz_contrib(res.mca, choice = "var", axes = 1, top = 15)
# Contributions of rows to dimension 2
fviz_contrib(res.mca, choice = "var", axes = 2, top = 15)
# Total contribution to dimension 1 and 2
fviz_contrib(res.mca, choice = "var", axes = 1:2, top = 15)
#we see that the the provider of the record,categories of the rate code are most contributive and the the type of the payment. Also it makes difference when it is night.
head(round(res.mca$var$contrib,2), 10)

plot.MCA(res.mca,choix=c("ind"),invisible=c("ind"),axes=c(3,4))
plot.MCA(res.mca,choix=c("var"),axes=c(3,4))

# Use modern ggplot facilities
names(res.mca)
fviz_eig(res.mca)
fviz_cos2(res.mca, choice = "var", axes = 1:2)+theme_bw()
fviz_contrib(res.mca, choice = "var", axes = 3:4)+theme_bw()
fviz_mca_var(res.mca, col.var="contrib",repel=TRUE)+
    scale_color_gradient2(low="green", mid="blue", 
    high="red", midpoint=0.75)+theme_bw()
fviz_mca_var(res.mca, col.var="contrib",repel=F)+
    scale_color_gradient2(low="green", mid="blue", 
    high="red", midpoint=0.75)+theme_bw()

fviz_mca_biplot(res.mca, invisible="ind",axes=1:2,repel=FALSE)+theme_bw()
#fviz_mca_biplot(res.mca, axes=1:2,repel=TRUE,arrows=c(TRUE,FALSE))+theme_bw()
```

# Synthesis through HCPC: Hierarchical Clustering

```{r}
###
### Clustering the individuals
### Before, hou have to perform a MCA with the number of axes 
### that you have decided to take into account (indicated through ncp=)
res.mca<-MCA(df[,c(vars_cat,"Total_amount")],quali.sup=c(5,13),quanti.sup=21,ncp=26)
res.hcpc<-HCPC(res.mca,nb.clust=8,order=TRUE)
names(res.hcpc)

fviz_cluster(res.hcpc, geom = "point", main = "Factor map")

#Anytip
vars_con_pca<-c(6,7,8,9,10,11,12,13,14,15,16,17,18,22,23,24,25,26) 
res.pca<-PCA(df[,vars_con_pca], quanti.sup = 13, quali.sup = 14, ncp = 6 ) # TotalAmount and AnyTip
res.hcpcPCA <-HCPC(res.pca, nb.clust = 8, order=TRUE)
table (res.hcpc$data.clust$clust, res.hcpcPCA$data.clust$clust)
claHPPCA<-factor(res.hcpcPCA$data.clust$clust,labels=paste("kHP-",1:8))
claHP<-factor(res.hcpc$data.clust$clust,levels=c(5,7,1,4,2,8,3,6),labels=c("kKM-5","kKM-7","kKM-1","kKM-4","kKM-2","kKM-8","kKM-3","kKM-6"))
tt<-table(claHPPCA,claHP)
tt
sum(diag(tt)/sum(tt))
#F.total
vars_con_pca<-c(6,7,8,9,10,11,12,13,14,15,16,17,18,23,24,25,26,41) 
names(df[,vars_con_pca])
names(df)
res.pca<-PCA(df[,vars_con_pca], quanti.sup = 13, quali.sup = 18, ncp = 6 ) # TotalAmount and f.total
res.hcpcPCA <-HCPC(res.pca, nb.clust = 8, order=TRUE)
claHPPCA<-factor(res.hcpcPCA$data.clust$clust,labels=paste("kHP-",1:8))
claHP<-factor(res.hcpc$data.clust$clust,levels=c(4,7,1,5,2,8,3,6),labels=c("kKM-5","kKM-7","kKM-1","kKM-4","kKM-2","kKM-8","kKM-3","kKM-6"))
tt<-table(claHPPCA,claHP)
tt
sum(diag(tt)/sum(tt))
```
