---
title: "New York cabs"
author: "Katerina Dimitrova, Jose Romero, Sergi Munoz"
date: "March 18, 2018"
output: pdf_document
---

```{r,include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)

```
```{r,include=FALSE}
setwd("D:/ADEI/ADEI.git/trunk") #Change 


```
#Previous work

## Load requiered packages

```{r,include=FALSE}
# Load Required Packages: to be increased over the course
requiredPackages <- c("effects","FactoMineR","car", "factoextra","RColorBrewer","ggplot2")
missingPackages <- requiredPackages[!(requiredPackages %in% installed.packages()[,"Package"])]

if(length(missingPackages)) install.packages(missingPackages)
lapply(requiredPackages, require, character.only = TRUE)
```

## Select 5000 samples
```{r}
#Load samples

#Use birthday of 1 member of the group
#set.seed(03101994)
#nrow(df)
#sam<-sample(1:nrow(df),5000)
#sam<-as.vector(sort(sam))

#df<-df[sam,]
#setwd("/Users/Sergi/Desktop/Sergi/CABS")
#df<-read.table("green_tripdata_2016-01.csv",header=T, sep=",")
#save.image("Taxi5000_raw.RData") # Dont execute again since it will create a new data and the following code must be ajusted..
```

##Load usefull functions

```{r}
countX <- function(x,X) {
  n_x <- NULL
  for (j in 1:ncol(x)) {n_x[j] <- sum(x[,j]==X) }
  n_x <- as.data.frame(n_x)
  rownames(n_x) <- names(x)
  nx_i <- rep(0,nrow(x))
  for (j in 1:ncol(x)) {nx_i <- nx_i + as.numeric(x[,j]==X) }
  list(nx_col=n_x,nx_ind=nx_i) }

countNA <- function(x) {
  mis_x <- NULL
  for (j in 1:ncol(x)) {mis_x[j] <- sum(is.na(x[,j])) }
  mis_x <- as.data.frame(mis_x)
  rownames(mis_x) <- names(x)
  mis_i <- rep(0,nrow(x))
  for (j in 1:ncol(x)) {mis_i <- mis_i + as.numeric(is.na(x[,j])) }
  list(mis_col=mis_x,mis_ind=mis_i) }

man.dist.manual <- function(p1Lat, p1Lon, p2Lat, p2Lon) {
  #return(abs(pointDistance(c(p1$lon, p1$lat), c(p1$lon, p2$lat), longlat=TRUE)) + abs(pointDistance(c(p2$lon, p1$lat), c(p2$lon, p2$lat), longlat=TRUE)))
  R = 6371
  lat1 = degrees.to.radians(p1Lat)
  lon1 = degrees.to.radians(p1Lon)
  lat2 = degrees.to.radians(p2Lat)
  lon2 = degrees.to.radians(p2Lon)
  A_lat = lat2 - lat1
  A_lon = lon2 - lon1
  a = sin(A_lat/2)^2
  c = 2 * atan2(sqrt(a), sqrt(1-a))
  dist_lat = R * c
  a = sin(A_lon/2)^2
  c = 2 * atan2(sqrt(a), sqrt(1-a))
  dist_lon = R * c
  abs(dist_lat) + abs(dist_lon)
  return(abs(dist_lat) + abs(dist_lon))
}

degrees.to.radians<-function(value) {
  return(value*0.0174532925)
}

```


## Delete unnecessary attributes
```{r}
load("Taxi5000_raw2.RData")
table(df$Ehail_fee) ##Delete unnecessary row
df$Ehail_fee<-NULL

```

#Conversion of qualitative variables

Numeric variables corresponding to qualitative concepts are converted to factors.
We also 

## VendorID
```{r}
#No missing Data
missingData<-which(is.na(df$VendorID));length(missingData) 
#No errors
errors<-which(df$VendorID==0.0);length(errors) 
df$VendorID<-factor(df$VendorID,labels=c("Creative Mobile Technologies, LLC","VeriFone Inc."))
table(df$VendorID)
barplot(prop.table(table(df$VendorID)))
```

## RateCodeID
```{r}
missingData<-which(is.na(df$RateCodeID));length(missingData) #No missing Data
errors<-which(df$RateCodeID==0.0);length(errors) #No errors
df$RateCodeID<-factor(df$RateCodeID,labels=c("Standard rate","JFK","Newark","Nassau or Westchester","Negotiated fare"))
table(df$RateCodeID) 
barplot(prop.table(table(df$RateCodeID)))
```

## Store_and_fwd_flag 
```{r}
missingData<-which(is.na(df$Store_and_fwd_flag));length(missingData) #No missing Data
erors<-which(df$Store_and_fwd_flag==0.0);length(errors) #No errors
df$Store_and_fwd_flag<-factor(df$Store_and_fwd_flag,labels=c("not a store and forward trip","store and forward trip"))
table(df$Store_and_fwd_flag)
barplot(prop.table(table(df$Store_and_fwd_flag)))
```

## Payment_type 
```{r}
missingData<-which(is.na(df$Trip_type));length(missingData) #No missing Data
errors<-which(df$Payment_type==0.0);length(errors) #No errors
df$Payment_type<-factor(df$Payment_type,labels=c("Credit card","Cash", "No charge", "Dispute"))
table(df$Payment_type)
barplot(prop.table(table(df$Payment_type)))
```

## Trip_type
```{r}
missingData<-which(is.na(df$Trip_type));length(missingData) #No missing Data
errors<-which(df$Trip_type==0.0);length(errors) #No errors
df$Trip_type<-factor(df$Trip_type,labels=c("Street-hail","Dispatch"))
table(df$Trip_type)
barplot(prop.table(table(df$Trip_type)))
```


# Univariant Descriptive Analysis 
## Passenger_count
```{r}
## Number of missing values:
missingData<-which(is.na(df$Passenger_count));length(missingData) #No missing Data
errors<-which(df$Passenger_count<=0.0);length(errors) #2 errors
outliers<-which(df$Passenger_count>6.0);length(outliers) #0 outlier
df[errors,"Passenger_count"]<-NA
df[outliers,"Passenger_count"]<-NA
boxplot(df$Passenger_count)
hist(df$Passenger_count, col="pink")
```

## Trip_distance
```{r} 
missingData<-which(is.na(df$Trip_distance));length(missingData) #No missing Data
errors<-which(df$Trip_distance<=0.0);length(errors) #59 errors

dfaux<-df
ll<-which(is.na(df$Trip_distance));ll
if(length(ll)>0){
  dfaux<-df[-ll,]
}
iqrvar<-IQR(dfaux$Trip_distance)
quantil3<-quantile(dfaux$Trip_distance, .75);quantil3 #get 3rd quartile
outliers<-which(df$Trip_distance>(iqrvar*3)+quantil3);length(outliers) #138 extreme outliers

df[outliers,"Trip_distance"]<-NA 
df[errors,"Trip_distance"]<-NA 
boxplot(df$Trip_distance)
hist(df$Trip_distance, col="pink")
```

## Pickup_longitude
```{r}
missingData<-which(is.na(df$Trip_distance));length(missingData) #No missing Data

#min and max longitudes for New York city boundaries
min_long <- -74.15 
max_long <- -73.7004

errors<-which(df$Pickup_longitude< min_long);length(errors)
errors<-c(errors,which(df$Pickup_longitude> max_long));length(errors)
errors<-c(errors,which(df$Pickup_longitude==0.0));length(errors)
df[errors,"Pickup_longitude"]<-NA #12 errors

ll<-which(is.na(df$Pickup_longitude));ll
if(length(ll)>0){
  dfaux<-df[-ll,]
}

iqrvar<-IQR(dfaux$Pickup_longitude)
quantil3<-quantile(dfaux$Pickup_longitude, .75);quantil3 #get 3rd quartile
quantil1<-quantile(dfaux$Pickup_longitude, .25);quantil1 #get 1st quartile

UpperOutlier<-which(df$Pickup_longitude>quantil3+(iqrvar*3));length(UpperOutlier) #14 extreme UpperOutliers
LowerOutlier<-which(df$Pickup_longitude<quantil1-(iqrvar*3));length(LowerOutlier) #1 extreme LowerOutliers


df[UpperOutlier,"Pickup_longitude"]<-NA 
df[LowerOutlier,"Pickup_longitude"]<-NA 
boxplot(df$Pickup_longitude)
summary(df$Pickup_longitude)
```

## Pickup_latitude
```{r}
missingData<-which(is.na(df$Pickup_latitude));length(missingData) #No missing Data

#we need to add here error control (what if longitude is out of scope?) and outlier management

summary(df$Pickup_latitude)

#min and max latitudes for New York city boundaries
min_lat <- 40.5774
max_lat <- 40.9176

errors<-which(df$Pickup_latitude< min_lat);length(errors)
errors<-c(errors,which(df$Pickup_latitude> max_lat));length(errors)
errors<-c(errors,which(df$Pickup_latitude==0.0));length(errors)
df[errors,"Pickup_latitude"]<-NA #17 errors

ll<-which(is.na(df$Pickup_latitude));ll
if(length(ll)>0){
  dfaux<-df[-ll,]
}

iqrvar<-IQR(dfaux$Pickup_latitude)
quantil3<-quantile(dfaux$Pickup_latitude, .75);quantil3 #get 3rd quartile
quantil1<-quantile(dfaux$Pickup_latitude, .25);quantil1 #get 1st quartile

UpperOutlier<-which(df$Pickup_latitude>quantil3+(iqrvar*3));length(UpperOutlier) #0 extreme UpperOutliers
LowerOutlier<-which(df$Pickup_latitude<quantil1-(iqrvar*3));length(LowerOutlier) #0 extreme LowerOutliers
df[UpperOutlier,"Pickup_latitude"]<-NA 
df[LowerOutlier,"Pickup_latitude"]<-NA 
boxplot(df$Pickup_latitude)
summary(df$Pickup_latitude)
```

##Dropoff_longitude
```{r}
missingData<-which(is.na(df$Dropoff_longitude));length(missingData) #No missing Data

errors<-c(errors,which(df$Dropoff_longitude==0.0));length(errors) #26 errors
df[errors,"Dropoff_longitude"]<-NA

ll<-which(is.na(df$Dropoff_longitude));ll
if(length(ll)>0){
  dfaux<-df[-ll,]
}

iqrvar<-IQR(dfaux$Dropoff_longitude)
quantil3<-quantile(dfaux$Dropoff_longitude, .75);quantil3 #get 3rd quartile
quantil1<-quantile(dfaux$Dropoff_longitude, .25);quantil1 #get 1st quartile

UpperOutlier<-which(df$Dropoff_longitude>quantil3+(iqrvar*3));length(UpperOutlier) #0 extreme UpperOutliers
LowerOutlier<-which(df$Dropoff_longitude<quantil1-(iqrvar*3));length(LowerOutlier) #0 extreme LowerOutliers
df[UpperOutlier,"Dropoff_longitude"]<-NA 
df[LowerOutlier,"Dropoff_longitude"]<-NA 

boxplot(df$Dropoff_longitude)
```

## Dropoff_latitude
```{r}
missingData<-which(is.na(df$Dropoff_latitude));length(missingData) #No missing Data

errors<-c(errors,which(df$Dropoff_latitude==0.0));length(errors) #35 errors
df[errors,"Dropoff_latitude"]<-NA

ll<-which(is.na(df$Dropoff_latitude));ll
if(length(ll)>0){
  dfaux<-df[-ll,]
}

iqrvar<-IQR(dfaux$Dropoff_latitude)
quantil3<-quantile(dfaux$Dropoff_latitude, .75);quantil3 #get 3rd quartile
quantil1<-quantile(dfaux$Dropoff_latitude, .25);quantil1 #get 1st quartile

UpperOutlier<-which(df$Dropoff_latitude>quantil3+(iqrvar*3));length(UpperOutlier) #0 extreme UpperOutliers
LowerOutlier<-which(df$Dropoff_latitude<quantil1-(iqrvar*3));length(LowerOutlier) #0 extreme LowerOutliers
df[UpperOutlier,"Dropoff_latitude"]<-NA 
df[LowerOutlier,"Dropoff_latitude"]<-NA 

boxplot(df$Dropoff_latitude)
```

## Fare_amount
```{r}
missingData<-which(is.na(df$Fare_amount));length(missingData) #No missing Data

sel<-which(df$Fare_amount<=0.0);length(sel) #10 missings
outlier<-which(df$Fare_amount>100);length(outlier) #1 outlier
df[sel,"Fare_amount"]<-NA 
df[outlier,"Fare_amount"]<-NA 
boxplot(df$Fare_amount)
hist(df$Fare_amount, col="pink")
summary(df$Fare_amount)

```
## Extra 
```{r}
missingData<-which(is.na(df$Extra));length(missingData) #No missing Data

sel<-which(df$Extra<0.0);length(sel) #10 missings
df[sel,"Extra"]<-NA 
boxplot(df$Extra)
hist(df$Extra, col="pink")
```

## MTA_tax
```{r}
missingData<-which(is.na(df$MTA_tax));length(missingData) #No missing Data

sel<-which(df$MTA_tax<0.0);length(sel) #103 missings
df[sel,"MTA_tax"]<-NA
boxplot(df$MTA_tax)
hist(df$MTA_tax, col="pink")
```

## Improvement_surcharge
```{r}
missingData<-which(is.na(df$improvement_surcharge));length(missingData) #No missing Data

sel<-which(df$improvement_surcharge<0.0);length(sel)
df[sel,"improvement_surcharge"]<-NA
boxplot(df$improvement_surcharge)
hist(df$improvement_surcharge, col="pink")
```

##Tip_amount
```{r}
missingData<-which(is.na(df$Tip_amount));length(missingData) #No missing Data

sel<-which(df$Tip_amount<0.0);length(sel) #107 missings
outlier<-which(df$Tip_amount>60.0);length(outlier) #1 missings
df[outlier,"Tip_amount"]<-NA
df[sel,"Tip_amount"]<-NA
boxplot(df$Tip_amount)
hist(df$Tip_amount, col="pink")
```

##Tolls_amount
```{r}
missingData<-which(is.na(df$Tolls_amount));length(missingData) #No missing Data

sel<-which(df$Tolls_amount<0.0);length(sel) #0 missings
df[sel,"Tolls_amount"]<-NA
boxplot(df$Tolls_amount)
hist(df$Tolls_amount, col="pink")
```

##Total_amount (Target)
```{r}
missingData<-which(is.na(df$Total_amount));length(missingData) #No missing Data

ll<-which(is.na(df$Total_amount));ll
if(length(ll)>0){
  dfaux<-df[-ll,]
}
iqrvar<-IQR(dfaux$Total_amount)
quantil3<-quantile(dfaux$Total_amount, .75) #get 3rd quartile
sel<-which(df$Total_amount<=0.0);length(sel) #22 errors
df[sel,"Total_amount"]<-NA 
outlier<-which(df$Total_amount>(iqrvar*3)+quantil3);length(outlier) #72 extreme outliers
df[outlier,"Total_amount"]<-NA 
boxplot(df$Total_amount)
hist(df$Total_amount, col="pink")
```

# Number of missing values:
```{r}

mis1<-countNA(df)
attributes(mis1)
rank(mis1$mis_col)
df$mis_ind <- mis1$mis_ind # new attribute missing values
mis1$mis_col

```

# Creating AnyTip
```{r}

df$AnyTip<-ifelse(df$Tip_amount<0.0001,0,1) 
df$AnyTip<-factor(df$AnyTip,labels=paste("AnyTip",c("No","Yes")))

```


#Creating auxiliar variables and doing their analysis

## Trip length
```{r}
for (i in 1:nrow(df)){
  df$trip_length[i] <-man.dist.manual(df$Pickup_latitude[i],df$Pickup_longitude[i],df$Dropoff_latitude[i],df$Dropoff_longitude[i])
}

summary(df$trip_length)
boxplot(df$trip_length)

```


## Trip distance in km
```{r}

df$trip_distance_km<-df$Trip_distance*1.609344 # Miles to km

summary(df$trip_distance_km)
boxplot(df$trip_distance_km)
```

## Travel time in minutes
```{r}

b1<-as.POSIXlt(df$lpep_pickup_datetime)
b2<-as.POSIXlt(df$Lpep_dropoff_datetime)
df$travel_time<-as.double(difftime(b2,b1,units='min'))
error<-which(df$travel_time== 0.0);length(error) #No missing Data
df[error,"travel_time"]<-NA 
summary(df$travel_time)
boxplot(df$travel_time)
```


## Pick_up_hour
```{r}
mydate <- as.POSIXlt(df$lpep_pickup_datetime)
df$pick_up_hour <- mydate$hour

summary(df$pick_up_hour)
boxplot(df$pick_up_hour)
```


## Pick_up_period
```{r}
# night, morning, valley and afternoon

df$pick_up_period= cut(df$pick_up_hour, breaks = c(-1, 5, 11, 17, 23), labels= c("night", "morning", "valley", "afternoon"))

summary(df$pick_up_period)

```


# Declaring vectors of data
```{r}
names(df)
vars_con<-names(df)[c(6,7,8,9,10,11,12,13,14,15,16,17,23,24,25,26)] 
vars_con
vars_dis<-names(df)[c(1,4,5,10,19,20,27,28)]
vars_dis
vars_res<-names(df)[c(18,22)]
vars_res

```


## Multivariant Outlier Detection
We haven't sorted some problems during implementation to achieve a logic result of aq.plot in order to identify multivariate outliers.

# Correlations error variable


```{r}

library(FactoMineR)
names (df)
summary(df$mis_ind)
corV <- cor(df[,vars_con], df$mis_ind,  use = "complete.obs")
corV 
 # rank 
rank(corV) 


```


# Imputation

## Remove observations with NA at targets

```{r}
ll<-which(is.na(df$Total_amount));
if(length(ll)>0){
  df<-df[-ll,]
}

ll<-which(is.na(df$AnyTip));ll
if(length(ll)>0){
  df<-df[-ll,]
}

```


## Imputation of numeric variables
```{r}

library(missMDA)
res.comp <- imputePCA(df[,vars_con], ncp=4)
summary(res.comp$completeObs)
df[,"Pickup_longitude"]<-res.comp$completeObs[,"Pickup_longitude"]
df[,"Pickup_latitude"]<-res.comp$completeObs[,"Pickup_latitude"]
df[,"Dropoff_longitude"]<-res.comp$completeObs[,"Dropoff_longitude"]
df[,"Dropoff_latitude"]<-res.comp$completeObs[,"Dropoff_latitude"]
df[,"Trip_distance"]<-res.comp$completeObs[,"Trip_distance"]
df[,"Passenger_count"]<-res.comp$completeObs[,"Passenger_count"]
df[,"Fare_amount"]<-res.comp$completeObs[,"Fare_amount"]
df[,"Extra"]<-res.comp$completeObs[,"Extra"]
df[,"travel_time"]<- res.comp$completeObs[,"travel_time"]
df[,"MTA_tax"]<-res.comp$completeObs[,"MTA_tax"]
df[,"Tip_amount"]<-res.comp$completeObs[,"Tip_amount"]
df[,"Tolls_amount"]<-res.comp$completeObs[,"Tolls_amount"]
df[,"improvement_surcharge"]<-res.comp$completeObs[,"improvement_surcharge"]
```

## Espeed (km/h)
```{r}
#efective speed : trigonometric distance between pickup point and dropoff point divided by travel time 

for (i in 1:nrow(df)){
  df$espeed[i] <- df$trip_length[i]/(df$travel_time[i]/60)
}
vars_con<-names(df)[c(6,7,8,9,10,11,12,13,14,15,16,17,23,24,25,26,28)] 

res.comp <- imputePCA(df[,vars_con], ncp=4)
df[,"espeed"]<-res.comp$completeObs[,"espeed"]

summary(df$espeed)
boxplot(df$espeed)
```


## Imputation of qualitative variables
```{r}
library(missMDA)
vars_dis_qual <-vars_dis[c(1,2,3,5,6)]
res.impute = imputeMCA(df[,vars_dis_qual],ncp = 3)
df[,"VendorID"]<-res.impute$completeObs[,"VendorID"]
df[,"RateCodeID"]<-res.impute$completeObs[,"RateCodeID"]
df[,"Payment_type"]<-res.impute$completeObs[,"Payment_type"]
df[,"Trip_type"]<-res.impute$completeObs[,"Trip_type"]

```


# Creating factors

## Factorize function:
```{r}
factorize<- function(x) {
  quantile(x,seq(0,1,0.1))
  pp<-quantile(x);pp
  breaks<-c(unique(pp))
  f.x<-factor(cut(x,breaks))
  return(f.x);
}
```

## f.passenger
```{r}

df$f.passenger<-factor(cut(df$Passenger_count,breaks=c(0,1,6)))
summary(df$f.passenger)

```

## f.distance
```{r}
df$f.distance<-factorize(df$Trip_distance) # NO VA be?
summary(df$f.distance) 
```

## f.pickup_longitude
```{r}

df$f.pickup_longitude<-factorize(df$Pickup_longitude)
summary(df$f.pickup_longitude)
```

## f.pickup_latitude
```{r}
df$f.pickup_latitude<-factorize(df$Pickup_latitude)
summary(df$f.pickup_latitude)
```

## f.dropoff_longitude
```{r}
df$f.dropoff_longitude<-factorize(df$Dropoff_longitude)
summary(df$f.dropoff_longitude)

```

## f.dropoff_latitude
```{r}
df$f.dropoff_latitude<-factorize(df$Pickup_latitude)
summary(df$f.dropoff_latitude) 
```

## f.fare_amount
```{r}
df$f.fare_amount<-factorize(df$Fare_amount)
summary(df$f.fare_amount)
```

## f.extra
```{r}
df$f.extra<-factorize(df$Extra)
summary(df$f.extra)
```

## f.MTA_tax
```{r}

df$f.MTA_tax<-factorize(df$MTA_tax)
summary(df$f.MTA_tax)

```

## f.Improvement_surcharge
```{r}
df$f.Improvement_surcharge<-factorize(df$improvement_surcharge)
summary(df$f.Improvement_surcharge)
```

## f.tip_amount
```{r}
df$f.tip_amount<-factor(df$Tip_amount)
summary(df$f.tip_amount)
```

## f.tolls_amount
```{r}
df$f.toll<-factor(cut(df$Tolls_amount,breaks=c(-1,1,50)))
summary(df$f.toll)
```

## f.total_amount
```{r}
df$f.total<-factorize(df$Total_amount)
summary(df$f.total)
```


# Profiling
```{r}
library(FactoMineR)
# Numeric Target Total_Amount
#vars_con;vars_dis
#names(df)

#condes(df,1)
#condes(df, num.var=18)

# Binary Target AnyTip
#vars_con;vars_dis
#names(df)

#catdes(df[,c(vars_dis,vars_con)],5)
```







