---
title: "CASE_STUDY "
author: "Katerina Dimitrova, Jose Romero, Sergi Muñoz"
date: "March 18, 2018"
output: pdf_document
---

```{r,include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)

```
```{r,include=FALSE}
setwd("/Users/Sergi/Desktop/Sergi/ADEI") #Change 


```
#Introduction

## Load requiered packages

```{r,include=FALSE}
# Load Required Packages: to be increased over the course

requiredPackages <- c("effects","FactoMineR","car", "factoextra","RColorBrewer","ggplot2")
missingPackages <- requiredPackages[!(requiredPackages %in% installed.packages()[,"Package"])]

if(length(missingPackages)) install.packages(missingPackages)
lapply(requiredPackages, require, character.only = TRUE)

```

## Select 5000 samples
```{r}
#Load samples

### Use birthday of 1 member of the group
set.seed(03101994)
nrow(df)
sam<-sample(1:nrow(df),5000)
sam<-as.vector(sort(sam))

df<-df[sam,]
#setwd("/Users/Sergi/Desktop/Sergi/CABS")
#df<-read.table("green_tripdata_2016-01.csv",header=T, sep=",")
#save.image("Taxi5000_raw.RData") # Dont execute again since it will create a new data and the following code must be ajusted..
```

##Load functions

```{r}
countX <- function(x,X) {
  n_x <- NULL
  for (j in 1:ncol(x)) {n_x[j] <- sum(x[,j]==X) }
  n_x <- as.data.frame(n_x)
  rownames(n_x) <- names(x)
  nx_i <- rep(0,nrow(x))
  for (j in 1:ncol(x)) {nx_i <- nx_i + as.numeric(x[,j]==X) }
  list(nx_col=n_x,nx_ind=nx_i) }

countNA <- function(x) {
  mis_x <- NULL
  for (j in 1:ncol(x)) {mis_x[j] <- sum(is.na(x[,j])) }
  mis_x <- as.data.frame(mis_x)
  rownames(mis_x) <- names(x)
  mis_i <- rep(0,nrow(x))
  for (j in 1:ncol(x)) {mis_i <- mis_i + as.numeric(is.na(x[,j])) }
  list(mis_col=mis_x,mis_ind=mis_i) }

man.dist.manual <- function(p1Lat, p1Lon, p2Lat, p2Lon) {
  #return(abs(pointDistance(c(p1$lon, p1$lat), c(p1$lon, p2$lat), longlat=TRUE)) + abs(pointDistance(c(p2$lon, p1$lat), c(p2$lon, p2$lat), longlat=TRUE)))
  R = 6371
  lat1 = degrees.to.radians(p1Lat)
  lon1 = degrees.to.radians(p1Lon)
  lat2 = degrees.to.radians(p2Lat)
  lon2 = degrees.to.radians(p2Lon)
  A_lat = lat2 - lat1
  A_lon = lon2 - lon1
  a = sin(A_lat/2)^2
  c = 2 * atan2(sqrt(a), sqrt(1-a))
  dist_lat = R * c
  a = sin(A_lon/2)^2
  c = 2 * atan2(sqrt(a), sqrt(1-a))
  dist_lon = R * c
  abs(dist_lat) + abs(dist_lon)
  return(abs(dist_lat) + abs(dist_lon))
}

degrees.to.radians<-function(value) {
  return(value*0.0174532925)
}

```


## Delete unnecessary attributes
```{r}
load("Taxi5000_raw2.RData")
table(df$Ehail_fee) ##Delete unnecessary row
df$Ehail_fee<-NULL

# Now one by one describe vars
names(df)
```

#Converting numeric variables corresponding to qualitative concepts to factors
## VendorID
```{r}
sel<-which(df$VendorID==0.0);length(sel) #No missing Data
df$VendorID<-factor(df$VendorID,labels=c("Creative Mobile Technologies, LLC","VeriFone Inc."))
summary(df$VendorID)
table(df$VendorID)
barplot(prop.table(table(df$VendorID)))
```

## RateCodeID, there was no group ride
```{r}
sel<-which(df$RateCodeID==0.0);length(sel) #No missing Data
df$RateCodeID<-factor(df$RateCodeID,labels=c("Standard rate","JFK","Newark","Nassau or Westchester","Negotiated fare"))
summary(df$RateCodeID)
table(df$RateCodeID)
barplot(prop.table(table(df$RateCodeID)))
```

## Store_and_fwd_flag 
```{r}
#//first the N and than Y
sel<-which(df$Store_and_fwd_flag==0.0);length(sel) #No missing Data
df$Store_and_fwd_flag<-factor(df$Store_and_fwd_flag,labels=c("not a store and forward trip","store and forward trip"))
summary(df$Store_and_fwd_flag)
table(df$Store_and_fwd_flag)
barplot(prop.table(table(df$Store_and_fwd_flag)))
```
## Payment_type 
```{r}
sel<-which(df$Payment_type==0.0);length(sel) #No missing Data
df$Payment_type<-factor(df$Payment_type,labels=c("Credit card","Cash", "No charge", "Dispute"))
summary(df$Payment_type)
table(df$Payment_type)
barplot(prop.table(table(df$Payment_type)))
```

## Trip_type
```{r}
sel<-which(df$Trip_type==0.0);length(sel) #No missing Data
df$Trip_type<-factor(df$Trip_type,labels=c("Street-hail","Dispatch"))
summary(df$Trip_type)
table(df$Trip_type)
barplot(prop.table(table(df$Trip_type)))
```


#Creating auxiliar variables
## Trip length

```{r}

for (i in 1:nrow(df)){
  df$trip_length[i] <-man.dist.manual(df$Pickup_latitude[i],df$Pickup_longitude[i],df$Dropoff_latitude[i],df$Dropoff_longitude[i])
}

```


## Trip distance in km
```{r}

df$trip_distance_km<-df$Trip_distance*1.609344 # Miles to km

```

## Travel time in minutes
```{r}

b1<-as.POSIXlt(df$lpep_pickup_datetime)
b2<-as.POSIXlt(df$Lpep_dropoff_datetime)
df$travel_time<-as.double(difftime(b2,b1,units='min'))

```

## Espeed (km/h)
```{r}
#efective speed : trigonometric distance between pickup point and dropoff point divided by travel time 

for (i in 1:nrow(df)){
  df$espeed[i] <- df$trip_length[i]/(df$travel_time[i]/60)
}

```

## Pick_up_hour
```{r}
mydate <- as.POSIXlt(df$lpep_pickup_datetime)
df$pick_up_hour <- mydate$hour

```


## Pick_up_period
```{r}
# night, morning, valley and afternoon

df$pick_up_period= cut(df$pick_up_hour, breaks = c(-1, 5, 11, 17, 23), labels= c("night", "morning", "valley", "afternoon"))


```



# Creating additional factors as a discretization 
## Factorize function:
```{r}
factorize<- function(x) {
  quantile(x,seq(0,1,0.1))
  pp<-quantile(x);pp
  breaks<-c(unique(pp))
  f.x<-factor(cut(x,breaks))
  return(f.x);
}
```

## Passenger_count
```{r}

## Number of missing values:
``{r}

df$f.passanger<-factor(cut(df$Passenger_count,breaks=c(0,1,6)))
summary(df$f.passanger) #2 NA
errors<-which(df$Passenger_count<=0.0);length(errors) #2 errors
outliers<-which(df$Passenger_count>6.0);length(outliers) #0 outlier
df[errors,"Passenger_count"]<-NA
df[outliers,"Passenger_count"]<-NA
boxplot(df$Passenger_count)
hist(df$Passenger_count, col="pink")
```

## Trip_distance
```{r}
df$f.distance<-factorize(df$Trip_distance) # NO VA be?
summary(df$f.distance) #59 missing values
errors<-which(df$Trip_distance<=0.0);length(errors) #59 errors

dfaux<-df
ll<-which(is.na(df$Trip_distance));ll
if(length(ll)>0){
  dfaux<-df[-ll,]
}
iqrvar<-IQR(dfaux$Trip_distance)
quantil3<-quantile(dfaux$Trip_distance, .75);quantil3 #get 3rd quartile
outliers<-which(df$Trip_distance>(iqrvar*3)+quantil3);length(outliers) #138 extreme outliers

df[outliers,"Trip_distance"]<-NA 
df[errors,"Trip_distance"]<-NA 
boxplot(df$Trip_distance)
hist(df$Trip_distance, col="pink")
```

## Pickup_longitude
```{r}

#min and max longitudes for New York city boundaries
min_long <- -74.15 
max_long <- -73.7004

df$f.longtitude<-factorize(df$Pickup_longitude)
summary(df$f.longtitude) #1 missing value <- this is not right

iqrvar<-IQR(df$Pickup_longitude)
quantil3<-quantile(df$Pickup_longitude, .75);quantil3 #get 3rd quartile
quantil1<-quantile(df$Pickup_longitude, .25);quantil1 #get 1st quartile

UpperOutlier<-which(df$Pickup_longitude>quantil3+(iqrvar*3));length(UpperOutlier) #20 extreme UpperOutliers
LowerOutlier<-which(df$Pickup_longitude<quantil1-(iqrvar*3));length(LowerOutlier) #0 extreme LowerOutliers

error<-which(df$Pickup_longitude< min_long);length(min_long) #1 error


sel<-which(df$Pickup_longitude==0.0 add df$Pickup_longitude<min_long || df$Pickup_longitude>max_long);length(sel) #11 errors
df[sel,"Pickup_longitude"]<-NA 
df[UpperOutlier,"Pickup_longitude"]<-NA 
df[LowerOutlier,"Pickup_longitude"]<-NA 
boxplot(df$Pickup_longitude)
summary(df$Pickup_longitude)
```

## Pickup_latitude
```{r}
#we need to add here error control (what if longitude is out of scope?) and outlier management

min_lat <- 40.5774
max_lat <- 40.9176


df$f.latitude<-factorize(df$Dropoff_latitude)
summary(df$f.latitude) #4 NAs
sel<-which(df$Pickup_latitude==0.0);length(sel) #11 missings
df[sel,"Pickup_latitude"]<-NA 
boxplot(df$Pickup_latitude)
hist(df$Pickup_latitude, col="pink")
```

##Dropoff_longitude
```{r}
#we need to add here error control (what if longitude is out of scope?) and outlier management

df$f.longtitudeDrop<-factorize(df$Dropoff_longitude)
summary(df$f.longtitudeDrop) # 1 NAs
sel<-which(df$Dropoff_longitude==0.0);length(sel) #11 missings
df[sel,"Dropoff_longitude"]<-NA 
boxplot(df$Dropoff_longitude)
```

## Dropoff_latitude
```{r}
#we need to add here error control (what if longitude is out of scope?) and outlier management

#quantile(df$Dropoff_latitude,seq(0,1,0.1))
#pp<-quantile(df$Dropoff_latitude);pp
#df$f.latitudeDrop<-factor(cut(df$Dropoff_latitude,pp))  # NO VA be
#summary(df$f.latitudeDrop) # 4 NAs ? Outlier
#boxplot(df$Pickup_latitude)
#hist(df$Pickup_latitude, col="pink")

min_lat <- 40.5774
max_lat <- 40.9176


df$f.latitudedeDrop<-factorize(df$Dropoff_latitude)
summary(df$f.latitudedeDrop) # 4 NAs
sel<-which(df$Dropoff_latitude==0.0);length(sel) #11 missings
df[sel,"Dropoff_latitude"]<-NA 
boxplot(df$Dropoff_latitude)
summary(df$Dropoff_latitude)
```

## Fare_amount
```{r}
df$f.fare_amount<-factorize(df$Fare_amount)
summary(df$f.fare_amount)

sel<-which(df$Fare_amount<=0.0);length(sel) #10 missings
outlier<-which(df$Fare_amount>100);length(outlier) #1 outlier
df[sel,"Fare_amount"]<-NA 
df[outlier,"Fare_amount"]<-NA 
boxplot(df$Fare_amount)
hist(df$Fare_amount, col="pink")
summary(df$Fare_amount)

```
## Extra 
```{r}
df$f.extra<-factorize(df$Extra)
summary(df$f.extra) #1 NA's

sel<-which(df$Extra<0.0);length(sel) #10 missings
df[sel,"Extra"]<-NA 
boxplot(df$Extra)
hist(df$Extra, col="pink")
```

## MTA_tax
```{r}
df$f.MTA_tax<-factorize(df$MTA_tax)
summary(df$f.MTA_tax) #11 NA's -> values of -0.5 => Outliers?
sel<-which(df$MTA_tax<0.0);length(sel) #103 missings
df[sel,"MTA_tax"]<-NA
boxplot(df$MTA_tax)
hist(df$MTA_tax, col="pink")
```

## Improvement_surcharge
```{r}
df$f.Improvement_surcharge<-factorize(df$improvement_surcharge)
summary(df$f.Improvement_surcharge) #11 NA's -> values of -0.3 => Outliers?

sel<-which(df$improvement_surcharge<0.0);length(sel) #107 missings
df[sel,"improvement_surcharge"]<-NA
boxplot(df$improvement_surcharge)
hist(df$improvement_surcharge, col="pink")
```

##Tip_amount
```{r}
df$f.tip_amount<-factor(df$Tip_amount)
summary(df$f.tip_amount) #2869 NA's
sel<-which(df$Tip_amount<0.0);length(sel) #107 missings
outlier<-which(df$Tip_amount>60.0);length(outlier) #1 missings
df[outlier,"Tip_amount"]<-NA
df[sel,"Tip_amount"]<-NA
boxplot(df$Tip_amount)
hist(df$Tip_amount, col="pink")
```

##Tolls_amount
```{r}
df$f.toll<-factor(cut(df$Tolls_amount,breaks=c(-1,1,50)))
summary(df$f.toll)
sel<-which(df$Tolls_amount<0.0);length(sel) #0 missings
df[sel,"Tolls_amount"]<-NA
boxplot(df$Tolls_amount)
hist(df$Tolls_amount, col="pink")
```

#Total_amount
```{r}
df$f.total<-factorize(df$Total_amount)
summary(df$f.total)

ll<-which(is.na(df$Total_amount));ll
if(length(ll)>0){
  dfaux<-df[-ll,]
}
iqrvar<-IQR(dfaux$Total_amount)
quantil3<-quantile(dfaux$Total_amount, .75) #get 3rd quartile
sel<-which(df$Total_amount<=0.0);length(sel) #22 errors
df[sel,"Total_amount"]<-NA 
outlier<-which(df$Total_amount>(iqrvar*3)+quantil3);length(outlier) #72 extreme outliers
df[outlier,"Total_amount"]<-NA 
boxplot(df$Total_amount)
hist(df$Total_amount, col="pink")
```
#Count per Variable:

## Number of missing values:
```{r}

mis1<-countNA(df)

attributes(mis1)
sort(mis1$mis_col)
df$mis_ind <- mis1$mis_ind # new attribute missing values
summary(mis1$mis_ind)
```

##Number of outliers ???
```{r}
outs<-rep(0,ncol(df))
show(outs)

```
### Multivariant Outlier Detection


#...In process
```{r}
#vars_con<-names(df)[c(6:9,11:18)] #Continuous variables


#install.packages('mvoutlier')
#library(mvoutlier) #not found??
#names(df)
#vars_con # Problems c(5,8,9,10,11,12)
#summary(df[,vars_con])
#vars_con_out<-vars_con[c(1:4)]
#aq.plot(df[,vars_con_out]) # Problems when few numeric values are present in one variable

# Use common sense, but technicalities might difficult the application of the procedure

#vars_con_out<-vars_con[c(1:4)]
#mvout<-aq.plot(df[,vars_con_out])  # Problems when missing data are present

# Use common sense
#vars_con
#vars_con_out<-vars_con[c(6,13,16)]
#aq.plot(df[,vars_con_out])  # Problems when missing data are present
#vars_con_out

```


```{r}
#install.packages("car")
#library(car) #not found??
#hist(df$Tip_amount, col="pink")
#catdes(data, 1)
```


# Imputation
## Definition of binary outcome: AnyTip
```{r}
# Binary Target: Any Tip?


df$AnyTip<-ifelse(df$Tip_amount<0.0001,0,1) 
df$AnyTip<-factor(df$AnyTip,labels=paste("AnyTip",c("No","Yes")))
#IMPORTANT
#if you touch some "global" variables you  will modify this part
# Now one by one describe vars
names(df)
vars_con<-names(df)[c(6,7,8,9,11,12,13,14,15,16,18)] 
vars_dis<-names(df)[c(1,4,10)]
vars_res<-names(df)[c(19,35)]
vars_res
vars_dis
```
## Imputation of numeric variables
```{r}
library(missMDA)
names(df)
vars_con_mis<-vars_con;length(vars_con_mis)
vars_con_mis
summary(df[,vars_con_mis])

res.comp <- imputePCA(df[,vars_con], ncp=4)
attributes(res.comp$completeObs)
summary(res.comp$completeObs)
df[,"Pickup_longitude"]<-res.comp$completeObs[,"Pickup_longitude"]
df[,"Pickup_latitude"]<-res.comp$completeObs[,"Pickup_latitude"]
df[,"Dropoff_longitude"]<-res.comp$completeObs[,"Dropoff_longitude"]
df[,"Dropoff_latitude"]<-res.comp$completeObs[,"Dropoff_latitude"]
df[,"Trip_distance"]<-res.comp$completeObs[,"Trip_distance"]
df[,"Fare_amount"]<-res.comp$completeObs[,"Fare_amount"]
df[,"Extra"]<-res.comp$completeObs[,"Extra"]
df[,"MTA_tax"]<-res.comp$completeObs[,"MTA_tax"]
df[,"Tip_amount"]<-res.comp$completeObs[,"Tip_amount"]
df[,"Tolls_amount"]<-res.comp$completeObs[,"Tolls_amount"]
df[,"Total_amount"]<-res.comp$completeObs[,"Total_amount"]
```

## Imputation of qualitative variables
```{r}
library(missMDA)
res.impute = imputeMCA(df[,vars_dis],ncp = 3)
res.impute
df[,"VendorID"]<-res.impute$completeObs[,"VendorID"]
#...necessaries variables
```
## Remove observations with NA at targets

```{r}
ll<-which(is.na(df$Total_amount));ll
if(length(ll)>0){
  df<-df[-ll,]
}

ll<-which(is.na(df$AnyTip));ll
if(length(ll)>0){
  df<-df[-ll,]
}

```

# Profiling
```{r}
#save(df,miss,vars_con,vars_dis,vars_res,file="MyTaxi5000_Clean.RData")
summary(df)

# Numeric Target Total_Amount
vars_con;vars_dis
names(df)

# condes(df[,c(vars_con,vars_dis)],1)
library(FactoMineR)
condes(df,15)

# Binary Target AnyTip
vars_con;vars_dis
names(df)

#catdes(df[,c(vars_dis,vars_con)],5)
catdes(df,which(names(df)=="AnyTip"))
```


```{r,include=FALSE}
library(rmarkdown)
#Generating the report from the Rmd file
#render("/home2/users/alumnes/1228816/ADEI/ADEI/Code_Doc.Rmd", "pdf_document") #execute this directly in the Rstuio console because here it will be recursive and will never end..
```





