---
title: "Session 9: NY Cabs Data - AnyTip Modelling - GLMz"
author: "Lidia Montero"
date: "May,25th 2018"
output: 
  html_document: 
    toc: true
    toc_depth: 3
    number_sections: true
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Introduction

## Data Description

Description http://www.nyc.gov/html/tlc/html/about/trip_record_data.shtml
Data Dictionary - SHL Trip Records -This data dictionary describes SHL trip data in visit http://www.nyc.gov/html/tlc/html/about/trip_record_data.shtml:

  - VendorID	A code indicating the LPEP provider that provided the record.      1= Creative Mobile Technologies, LLC; 2= VeriFone Inc.   
  - lpep_pickup_datetime	The date and time when the meter was engaged.    
  - lpep_dropoff_datetime	The date and time when the meter was disengaged.
  - Passenger_count	The number of passengers in the vehicle. This is a driver-entered value.    
  -  Trip_distance 	The elapsed trip distance in miles reported by the taximeter.   
  - Pickup_longitude	 Longitude where the meter was engaged.   
  - Pickup_latitude	Latitude where the meter was engaged.   
  - RateCodeID	The final rate code in effect at the end of the trip.  1= Standard rate  2=JFK 3=Newark 4=Nassau or Westchester 5=Negotiated fare 6=Group ride   
  - Store_and_fwd_flag	This flag indicates whether the trip record was held in vehicle memory before sending to the vendor, aka "store and forward," because the vehicle did not have a connection to the server: Y= store and forward trip  N= not a store and forward trip   
  - Dropoff_longitude	Longitude where the meter was timed off.   
  - Dropoff_ latitude	Latitude where the meter was timed off.   
  - Payment_type	A numeric code signifying how the passenger paid for the trip.  1= Credit card 2= Cash 3= No charge 4= Dispute 5= Unknown 6= Voided trip   
  - Fare_amount	The time-and-distance fare calculated by the meter.   
  - Extra	 Miscellaneous extras and surcharges.  Currently, this only includes the $0.50 and $1 rush hour and overnight charges. 
  - MTA_tax	 $0.50 MTA tax that is automatically triggered dfd on the metered rate in use.    - Improvement_surcharge	$0.30 improvement surcharge assessed on hailed trips at the flag   drop. The improvement surcharge began being levied in 2015.   
  - Tip_amount	 This field is automatically populated for credit card tips. Cash tips are not included.   
  - Tolls_amount	Total amount of all tolls paid in trip.    
  - Total_amount	The total amount charged to passengers. Does not include cash tips.   
  - Trip_type	A code indicating whether the trip was a street-hail or a dispatch that is automatically assigned dfd on the metered rate in use but can be altered by the driver. 
1= Street-hail 2= Dispatch  


## Load Required Packages

```{r, echo=FALSE, include=FALSE}
rm(list=ls())
# Load Required Packages: to be increased over the course

requiredPackages <- c("effects","FactoMineR","car", "factoextra","RColorBrewer","ggplot2","missMDA","mvoutlier","dplyr","ggmap","ggthemes","knitr","MVA")
missingPackages <- requiredPackages[!(requiredPackages %in% installed.packages()[,"Package"])]

if(length(missingPackages)) install.packages(missingPackages)
lapply(requiredPackages, require, character.only = TRUE)


```
# Statistical Modelling

## Load your sample after data cleaning and validation (Deliverable 1)


```{r}
#setwd("E:/DOCENCIA/FIB-ADEI/PRACTICA/NYCABS/LABS") # Choose your own directory
#file_path <- "E:/DOCENCIA/FIB-ADEI/PRACTICA/NYCABS/LABS/" 
#load(paste0(file_path,"MyTaxi5000_Modeling.RData"))
names(df)
summary(df)
vars_con
vars_dis
vars_res

m1<-glm(AnyTip~poly(trip_distance_km,2),data=df,family=binomial)

maux<-glm(AnyTip~Total_amount+Dropoff_longitude+,data=df,family=binomial)

m1b<-glm(AnyTip~trip_distance_km+I(trip_distance_km^2),data=df,family=binomial)
summary(m1b)

Anova(m1,test="Wald")
marginalModelPlot(m1)


# Quines?
names(df)
catdes(df,num.var=22)


# Feature Selection 
table(df$AnyTip,df$Payment_type)


```

# Generalized Linear Regression issues

*Target binary factor AnyTip*
```{r}
# Split your sample: work and test
ll<-which(df$Payment_type=="Cash");length(ll)
df<-df[-ll,]

set.seed(1234)
llwork<-sample(1:nrow(df),0.70*nrow(df),replace=FALSE)
llwork<-sort(llwork);length(llwork)
dfwork<-df[llwork,]
dftest<-df[-llwork,]

table(df$Extra)
table(df$MTA_tax)
df$f.extra<-0
df$f.extra[df$Extra>0]<-1
df$f.extra[df$Extra>0.5]<-2
df$f.extra<-factor(df$f.extra,labels=c("Extra-No","Extra-0.5","Extra-1"))
df$f.tax<-ifelse(df$MTA_tax>0,1,0)
df$f.tax<-factor(df$f.tax,labels=c("TAX-No","TAX-Yes"))

vars_con
catdes(df[,c(vars_con,"AnyTip")],num.var=18)

names(df)
vars_cexp<-vars_con[c(1,3,9,12,8)]
#vars_cexp<-df[c(18,8,6,24)]
```

# Modelling AnyTip using numeric explanatory variables

```{r}
m50<-glm(AnyTip~.,family="binomial",data=dfwork[,c("AnyTip",vars_cexp)])
summary(m50)
Anova(m50,test="Wald")
vif(m50)

vars_con
vars_cexp<-vars_con[c(1,3,9,12,8,15:17)]

m50<-glm(AnyTip~.,family="binomial",data=dffwork[,c("AnyTip",vars_cexp)])
summary(m50)
Anova(m50,test="Wald")
vif(m50)


#Después de molta feina... m73
m73<-m51

```


#Diagnostics
```{r}
#Boxplot dels residus
Boxplot(rstudent(m73), id.n=15)
llout<-which(abs(rstudent(m73))>2);length(llout)
dfwork[llout,]
```


